--- 
title: "DNA Damage Estimation"
author: "Peter Diakumis"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Estimation of DNA Damage in WGS Samples"
---

# Introduction {-}

Here we'll use the DNA Damage Estimator (DDE) toolkit
developed by Laurence Ettwiller (https://github.com/Ettwiller/Damage-estimator and
http://science.sciencemag.org/content/355/6326/752.long) to
estimate sample damage for a set of WGS tumor samples.
We are interested in running this tool on all the UMCCR samples processed with
bcbio, so it is important that we keep the runtime low.

* Section \@ref(dde-summary) has a detailed summary of the DDE functionality
* Section \@ref(dde-issues) has a few simple ways we can save computational
  time when running DDE

#### Load R Packages {-}

```{r knitr_opts, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r load_pkgs, message=FALSE, warning=FALSE}
require(tidyverse)
```


